
%% \criteriaDoc
%%    1 - criteria name without punctuation
%%    2 - criteria name with punctuation
%%    3 - syntax
%%    4 - applicable to
%%    5 - blurb
%%    6 - additional text --- wrap in \hasExpl, or \noExpl
%%    7 - example type --- \noEx, \singleEx, \multiEx, \tabbingEx
%%    8 - example(s)

%% \kwdCriteriaDoc
%%    1 - criteria name without colon
%%    2 - syntax
%%    3 - applicable to
%%    4 - blurb
%%    5 - additional text --- wrap in \hasExpl, or \noExpl
%%    6 - example type --- \noEx, \singleEx, \multiEx, \tabbingEx
%%    7 - example(s)

\def\kwdCriteriaDoc#1#2#3#4#5#6#7{\criteriaDoc{#1}{:#1}{#2}{#3}{#4}{#5}{#6}{#7}}

\begin{criteriaGroup}{Basic criteria}{}

\kwdCriteriaDoc{true}{:true}
{The form is evaluated at testing time; the criterion requires the result to be non-nil.}
{A single form.}
{\noExpl}
{\noEx}{}

\kwdCriteriaDoc{eq}{(:eq FORM)}
{The criterion argument and the form under test are both evaluated at testing time; the criterion requires that the results be \texttt{eq}.}
{A single form.}
{\noExpl}
{\singleEx}{(def-test eq1 (:eq 'b) (cadr '(a b c)))}

\kwdCriteriaDoc{symbol}{(:symbol NAME)}
{The form under test is evaluated at testing time.  The criterion requires that the result be a symbol which is \texttt{eq} to the symbol name given as the criterion argument.}
{A single form.}
{\noExpl}
{\multiEx}
{\passExample{(def-test sym1  (:symbol a) (car '(a b c)))}
 \failExample{(def-test sym1x (:symbol a) (cadr '(a b c)))}}

\kwdCriteriaDoc{eql}{(:eql FORM)}
{The criterion argument and the form under test are both evaluated at testing time; the criterion requires that the results be \texttt{eql}.}
{A single form.}
{\noExpl}
{\singleEx}{(def-test eql1 (:eql 2) (cadr '(1 2 3)))}

\kwdCriteriaDoc{equal}{(:equal FORM)}
{The criterion argument and the form under test are both evaluated at testing time; the criterion requires that the results be \texttt{equal}.}
{A single form.}
{\noExpl}
{\noEx}{}

\kwdCriteriaDoc{equalp}{(:equalp FORM)}
{The criterion argument and the form under test are both evaluated at testing time; the criterion requires that the results be \texttt{equalp}.}
{A single form.}
{\noExpl}
{\noEx}{}

\kwdCriteriaDoc{forms-eq}{:forms-eq}
{The two forms under test are both evaluated at testing time; the criterion requires that the results be \texttt{eq}.}
{Exactly two forms.}
{\noExpl}
{\singleEx}{(def-test eqforms1 :forms-eq (cadr '(a b c)) (caddr '(a c b)))}

\kwdCriteriaDoc{forms-eql}{:forms-eql}
{The two forms under test are both evaluated at testing time; the criterion requires that the results be \texttt{eql}.}
{Exactly two forms.}
{\noExpl}
{\singleEx}{(def-test eqlforms1 :forms-eql (cadr '(a 3 c)) (caddr '(a c 3)))}

\kwdCriteriaDoc{forms-equal}{:forms-equal}
{The two forms under test are both evaluated at testing time; the criterion requires that the results be \texttt{equal}.}
{Exactly two forms.}
{\noExpl}
{\noEx}{}

\kwdCriteriaDoc{predicate}{(:predicate FUNCTION-FORM)}
{The criterion argument is a symbol (unquoted) or a lambda expression;
  at testing time, the forms under test are evaluated and passed to
  the denoted function.  The criterion expects that the result of the
  function is non-nil.}
{Forms matching the input lambda list of the \texttt{FUNCTION-FORM}.}
{\noExpl}
{\multiEx}
{\passExample{(def-test pred1 (:predicate numberp) 3)}
 \passExample{(def-test pred2 (:predicate eql) (+ 1 2) 3)}}

\kwdCriteriaDoc{err}{(:err [:type CLASS])}
{At testing time, evaluates the form under test, expecting the
  evaluation to raise some condition.  If the \textit{CLASS} argument
  is supplied, the criterion expects the raised condition to be a
  subclass.  Note that the name of the type should \emph{not} be
  quoted; it is not evaluated.}
{Any.}
{\noExpl}
{\multiEx}
{\passExample{(def-test err1 (:err :type error) (error "this should be caught"))}
 \passExample{(def-test err2 (:err) (error "this should be caught"))}}

\kwdCriteriaDoc{perf}{(:perf [ :ns | :sec | :min ] TIME)}
{Evaluates the forms under test at testing time, and expects the
  evaluation to complete within the given time limit.}
{Any.}
{\noExpl}
{\singleEx}{(def-test perf1 (:perf :min 2) (ack 3 5))}

\end{criteriaGroup}

\begin{criteriaGroup}{Compound criteria}{}

\kwdCriteriaDoc{not}{(:not CRITERION)}
{Passes when testing according to \texttt{CRITERION} fails (but
does not throw an error).}
{As required by the subordinate criterion.}
{\noExpl}
{\singleEx}{(def-test not1 (:not (:symbol b)) 'a)}

\kwdCriteriaDoc{all}{(:all CRITERION CRITERION ...\ CRITERION)}
{This criterion brings several other criteria under one check, and
verifies that they all pass.}
{As required by the subordinate criteria.}
{\noExpl}
{\tabbingEx}{\texttt{(de}\=\texttt{f-}\=\texttt{check not1 ()}
\\ \>\>\texttt{(:all }\=\texttt{(:predicate even-p)}
\\ \>\>\>\texttt{(:predicate prime-p))}
\\ \>\texttt{2)}}

\kwdCriteriaDoc{any}{(:any CRITERION CRITERION ...\ CRITERION)}
{Passes when any of the subordinate criteria pass.}
{As required by the subordinate criteria.}
{\noExpl}
{\tabbingEx}{
\texttt{(de}\=\texttt{f-}\=\texttt{check not1 ()}
\\ \>\>\texttt{(:any }\=\texttt{(:predicate even-p)}
\\ \>\>\>\texttt{(:predicate prime-p))}
\\ \>\texttt{5)}}

\kwdCriteriaDoc{apply}{(:apply FUNCTION CRITERION)}
{At testing time, first evaluates the forms under test, applying
  \texttt{FUNCTION} to them.  The overall criterion passes or fails
  exactly when the subordinate \texttt{CRITERION} with the
  application's multiple result values.}
{Forms matching the input lambda list of the \texttt{FUNCTION}.}
{\noExpl}
{\singleEx}{(def-test applycheck (:apply cadr (:eql 10)) '(0 10 20))}

\kwdCriteriaDoc{check-err}{(:check-err CRITERION)}
{Like \texttt{:err}, but proceeds according to the subordinate
criterion rather than simply evaluating the input forms.}
{As required by the subordinate criterion.}
{\hasExpl{The difference between \texttt{:check-err} and \texttt{:err}
is that the latter deals only with evaluation of a form, whereas
\texttt{:check-err} is more about the unit testing process.  This form
is mostly useful for temporarily disregarding certain checks until
some later fix, when they \emph{won't} throw an error.}}
{\tabbingEx}
{\texttt{(de}\=\texttt{f-}\=\texttt{test check-err1}
\\ \> \> \texttt{(:check-err :forms-eq)}
\\ \> \texttt{'asdfgh (error "this should be caught"))}}

\kwdCriteriaDoc{progn}{(:progn FORM FORM ...\ FORM CRITERION)}
{At testing time, first evaluates the \texttt{FORM}s in order, and
  then proceeds with evaluation of the forms under test according to
  the subordinate criterion.}
{As required by the subordinate criterion.}
{\noExpl}
{\singleEx}
{(def-test form1 (:progn (setf zz 3) (:eql 3)) zz)}

\kwdCriteriaDoc{proj}{(:proj (INDEX INDEX ...\ INDEX) CRITERION)}
{Rearranges the forms under test by selecting a new list according to
  the index numbers into the old list.  Checking of the reorganized
  forms continues according to the subordinate criterion.}
{At least as many as to be accessible to the largest index.}
{\hasExpl{Note that containing criteria may have reordered forms and
    value from the original check.}}
{\tabbingEx}
{\texttt{(d}\=\texttt{ef}\=\texttt{-test proj-1}
\\ \>\> \texttt{(:proj (0 2) :forms-eq)}
\\ \> \texttt{'a 3 (car '(a b)))}}

\end{criteriaGroup}

\begin{criteriaGroup}{Criteria for multiple values}{
NST's approach to multiple values is stricter than Common Lisp's view
in the language itself.  In Lisp programs, additional returned values
are eminently ignorable; in fact some extra programming overhead is
required to access them.  However, in designing NST we draw a
distinction between \emph{using} a function on the one hand, and
\emph{designing} and \emph{testing} it on the other.  For example,
when using the \texttt{floor} function one often needs only the
quotient, and it is correspondingly easy to ignore the additional
argument which gives the remainder.  However, for the implementor of
\texttt{floor} it is impormtant that both results be predictable and
verified.\vspace{\parskip}\\\hspace*{\parindent}NST encourages the
thoughtful, consistent design and correct implementation of functions
returning multiple values by enforcing that tested forms generating
multiple variables should be paired with criteria for multiple values.
\emph{Any mismatch between the quantity of values returned by an
  evaluation and the quantity of values expected by a criterion is
  interpreted as a test failure}.  Dually to the usage mode of Common
Lisp, NST requires the additional overhead of the
\texttt{:drop-values} criterion to simply ignore additional returned
values.}

\kwdCriteriaDoc{value-list}{(:value-list CRITERION)}
{Converts multiple values into a single list value.}
{Arbitrarily many values.}
{\noExpl}{\noEx}{}

\kwdCriteriaDoc{values}{(:values CRITERION CRITERION ...\ CRITERION)}
{Checks each of the forms under test according to the respective
  subordinate criterion.}
{Exactly as many forms as subordinate criteria.}
{\noExpl}{\noEx}{}

\kwdCriteriaDoc{drop-values}{(:drop-values CRITERION)}
{Checks the primary value according to the subordinate criterion,
  ignoring any additional returned values from the evaluation of the
  form under test.}
{Any.}
{\noExpl}{\noEx}{}

\end{criteriaGroup}

\begin{criteriaGroup}{Criteria for lists}{}

\kwdCriteriaDoc{each}{(:each CRITERION)}
{At testing time, evaluates the form under test, expecting to find a
  list as a result.  Expects that each argument of the list according
  to the subordinate \texttt{CRITERION}, and passes when all of these
  checks pass.}
{A single form which evaluates to a list.}
{\noExpl}
{\singleEx}
{(def-test each1 (:each (:symbol a)) '(a a a a a))}

\kwdCriteriaDoc{seq}{(:seq CRITERION CRITERION ...\ CRITERION)}
{Evaluates its input form, checks each of its elements according to
  the respective subordinate criterion, and passes when all of them
  pass.}
{A single form which evaluates to a list with the same number of
elements as there are subordinate criteria to the
\texttt{:seq}.}
{\hasExpl{Note that \texttt{:seq} expects that the length of the list
will be the same as the number of subordinate criteria, and will fail
otherwise.}}
{\tabbingEx}{
(de\=f-\=check seqcheck
\\ \>\>  (:seq (:predicate symbolp) (:eql 1) (:symbol d))
\\ \>  '(a 1 d))
}

\kwdCriteriaDoc{permute}{(:permute CRITERION)}
{At testing time, evaluates the form under test, expecting to find a
  list as a result.  The criterion expects to find that some
  permutation of this list will satisfy the subordinate criterion.}
{A single form evaluating to a list.}
{\noExpl}{\tabbingEx}{
\textrm{Examples:}\ \=(def-test permute1 (:permute (:each (:eq 'a))) '(a a))
\\ \> (de\=f-\=check permute2
\\ \> \> \> (:permute (:seq \=(:symbol b)
\\ \> \> \>             \>(:predicate symbolp)
\\ \> \> \>             \>(:predicate numberp)))
\\ \> \>'(1 a b))
}

\end{criteriaGroup}

\begin{criteriaGroup}{Criteria for vectors}{}

\kwdCriteriaDoc{across}{(:across CRITERION CRITERION ...\ CRITERION)}
{Like \texttt{:seq}, but for a vector instead of a list.}
{A single form evaluating to a vector.}
{\noExpl}{\tabbingEx}{
(de\=f-\=check across1
\\ \>\>  (:across (:predicate symbolp) (:eql 1))
\\ \>  (vector 'a 1))}

\end{criteriaGroup}

\begin{criteriaGroup}{Criteria for classes}{}

\kwdCriteriaDoc{slots}{\small(:slots (NAME CRT) (NAME CRT) ...\ (NAME
    CRT))}
{Evaluates its input form, and passes when the value at each given
slot satisfies the corresponding subordinate constraint.}
{A single form evaluating to a class or struct object.}
{\hasExpl{Use of this criterion with structs rather than classes does
work on many platforms, since the CL specification defines
\texttt{with-slots} on classes only.}}
{\tabbingEx}{
(defc\=lass classcheck ()
\\ \> (\=(s1 :initarg :s1 :reader get-s1)
\\ \> \> (s2 :initarg :s2)
\\ \> \> (s3 :initarg :s3)))
\\ (def-test \=slot1
\\ \>  (:slots \=(s1 (:eql 10))
\\ \>  \>(s2 (:symbol zz))
\\ \>  \>(s3 (:seq \=(:symbol q) (:symbol w)
\\ \>  \> \> (:symbol e) (:symbol r))))
\\ \>  (m\=ake-instance 'classcheck
\\ \> \>  :s1 10 :s2 'zz :s3 '(q w e r)))
}

\end{criteriaGroup}

\begin{criteriaGroup}{Special criteria}{}

\kwdCriteriaDoc{sample}{(:sample \&key domains where verify values sample-size qualifying-sample max-tries)}
{Experimentally test a program property by generating random data.  See \manualOrRef{Section~\ref{quickcheck}}{the users' manual} for more information.}
{Forms must match lambda list \texttt{values}.}
{\noExpl}{\noEx}{}

\end{criteriaGroup}

\begin{criteriaGroup}{Programmatic and debugging criteria}{}

\kwdCriteriaDoc{info}{(:info MESSAGE SUBCRITERION)}
{Add an informational note to the check result.}
{Any.}
{\noExpl}
{\singleEx}{(def-test known-bug (:info "Known bug" (:eql 3)) 4)}

\kwdCriteriaDoc{pass}{:pass}
{A trivial test, which always passes.}
{Any.}
{\noExpl}
{\singleEx}{(def-test passing-test :pass 3 4 "sd")}

\kwdCriteriaDoc{fail}{(:fail FORMAT ARG ...\ ARG)}
{A trivial test, which always fails.  The format string and arguments
should be suitable for the Lisp \texttt{format} function.}
{Any.}
{\noExpl}
{\singleEx}{(def-test fails (:fail "Expected a \~{}a" "string") 312)}

\kwdCriteriaDoc{warn}{(:warn FORMAT ARG ...\ ARG)}
{Issue a warning.  The format string and arguments should be suitable for
 the Lisp \texttt{format} function.}
{Any.}
{\noExpl}
{\singleEx}{(:warn "\~{}d is not a perfect square" 5)}

\kwdCriteriaDoc{dump-forms}{(:dump-forms FORMAT)}
{For debugging NST criteria: fails after writes the current forms to
standard output.}
{Arbitrarily many values, compatible with the given string for the
Lisp \texttt{format} function.}
{\noExpl}{\noEx}{}

\end{criteriaGroup}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../manual/manual"
%%% TeX-PDF-mode: t
%%% End: 
